Query Optimization for Dynamic Imputation

### 2017-03-01 paper/examples.tex
paper: minor fixes

Before she can perform her queries, the epidemiologist must develop a strategy
for handling missing values. She currently has two options:
1) she could drop records with missing values in relevant fields,
2) she could use a standard data imputation package on all of her data. Both of these
approaches have significant drawbacks. For the query pictured in~\Cref{fig:example-query}, 
(1) drops 1492 potentially relevant tuples,
while her previous experiences with (2) have proven to be expensive. The epidemiologist needs a more complete picture
of the data, so (1) is insufficient, and she is under time pressure, so (2)
is infeasible.

>>>>

Before she can perform her queries, the epidemiologist must develop a strategy
for handling missing values. She currently has two options:
(1) she could drop records with missing values in relevant fields,
(2) she could use a standard data imputation package on all of her data. Both of these
approaches have significant drawbacks. For the query pictured in~\Cref{fig:example-query}, 
(1) drops 1492 potentially relevant tuples,
while her previous experiences with (2) have proven to be expensive. The epidemiologist needs a more complete picture
of the data, so (1) is insufficient, and she is under time pressure, so (2)
is infeasible.

### 2017-03-01 paper/introduction.tex
paper: minor fixes

% Solution
One common approach is to manually replace missing values using some sort of statistical or predictive model based on other values in the table and record.
This process is called \emph{imputation}.
In this paper, we present a new system, called \ProjectName{}, which is designed to selectively apply imputation to a subset of records, on-the-fly, during query execution.  \emph{The key insight behind \ProjectName{} is that imputation
only needs to be performed on the data relevant to a particular query and
that this subset is generally much smaller than the entire database.}  
% Background
While traditional imputation methods work over the entire data set and replace all missing
values, running a sophisticated imputation algorithm over a large data set can be very
expensive---our experiments show that even a relatively simple decision tree algorithm takes \acsbaseresulthours{} to train and run
on a 600K row database.

@@

\subsection{Contributions}
\ProjectName{} leverages a number of contributions to minimize imputation time while producing comparable results to traditional approaches,
increasing productivity for users who frequently encounter missing values in their data. These contributions include:
\begin{itemize}
\item \textbf{Relational algebra extended with imputation:}
  We extend the standard relational algebra with two new operators to represent imputation operations: impute ($\mu$) and drop ($\delta$).
  The impute operation fills in missing data values using any statistical imputation
  technique, such as chained-equation decision trees ~\cite{burgette2010multiple}.
  The drop operation simply drops tuples which contain \nullv{} values.
\item \textbf{Model of imputation quality and cost:}
  We extend the traditional cost model for query plans to incorporate a measure of the quality of the imputations performed.
  We use the cost model to abstract over the imputation algorithm used.
  To add an imputation technique, it is sufficient to characterize it with two functions: one to describe its running time and one to describe the quality of its results.
\item \textbf{Query planning with imputation:}
  We present the first query planning algorithm to jointly optimize for running time and the quality of the imputed results.
  It does so by maintaining multiple sets of Pareto optimal plans according to the cost model.
  By deferring selection of the final plan, we make it possible for the user to trade off running time and result quality.
%\item \textbf{Proof of optimality:}
% We prove that our cost-model and planning algorithm produce a sound and complete final Pareto frontier, making the final plan choice optimal under our search space and the user's preferences.
\end{itemize}

>>>>

% Solution
One common approach is to manually replace missing values using some sort of statistical or predictive model based on other values in the table and record.
This process is called \emph{imputation}.
In this paper, we present a new system, called \ProjectName{}, which is designed to selectively apply imputation to a subset of records, on-the-fly, during query execution.  \emph{The key insight behind \ProjectName{} is that imputation
only needs to be performed on the data relevant to a particular query and
that this subset is generally much smaller than the entire database.}  
% Background
While traditional imputation methods work over the entire data set and replace all missing values, running a sophisticated imputation algorithm over a large data set can be very expensive---our experiments show that even a relatively simple regression tree algorithm takes \acsbaseresulthours{} to train and run
on a 600K row database.

@@

\subsection{Contributions}
\ProjectName{} leverages a number of contributions to minimize imputation time while producing comparable results to traditional approaches.
These contributions include:
\begin{itemize}
\item \textbf{Relational algebra extended with imputation:}
  We extend the standard relational algebra with two new operators to represent imputation operations: impute ($\mu$) and drop ($\delta$).
  The impute operation fills in missing data values using any statistical imputation
  technique, such as chained-equation decision trees ~\cite{burgette2010multiple}.
  The drop operation simply drops tuples which contain \nullv{} values.
\item \textbf{Model of imputation quality and cost:}
  We extend the traditional cost model for query plans to incorporate a measure of the quality of the imputations performed.
  We use the cost model to abstract over the imputation algorithm used.
  To add an imputation technique, it is sufficient to characterize it with two functions: one to describe its running time and one to describe the quality of its results.
\item \textbf{Query planning with imputation:}
  We present the first query planning algorithm to jointly optimize for running time and the quality of the imputed results.
  It does so by maintaining multiple sets of Pareto optimal plans according to the cost model.
  By deferring selection of the final plan, we make it possible for the user to trade off running time and result quality.
%\item \textbf{Proof of optimality:}
% We prove that our cost-model and planning algorithm produce a sound and complete final Pareto frontier, making the final plan choice optimal under our search space and the user's preferences.
\end{itemize}
